Peptide retention time prediction

### 2014-11-09 retention.tex
Added paragraphs about mf

@@

% what is the idea 
%In Paper IV we tested this hypothesis by calculating and evaluating
%two types of nonlinear gradients. Our results showed that the new
%gradients not only provided more uniform distributions of the
%peptides, but also led to an increase in the number of peptide
%identifications. In addition, they facilitated the identification of
%different peptide species compared to the typical linear gradients.
%The project was carried out in collaboration with the Mechtler group
%at the Research Institute of Molecular Pathology, Vienna.

>>>>

In our group we have also investigated the possibility to use predicted retention times to increase the throughput of proteomics experiments. Our starting point was the observation that in typical shotgun proteomics experiments only a small fraction of the detected ions is
fragmented, and thus used to identify
peptides. The remaining unfragmented ions are usually disregarded in
the subsequent data analyses. In the light of these observations, we investigated whether we could use
information encapsulated by unfragmented ions to improve the protein
inference in shotgun proteomics \cite{mf}.

% how it works and main results
More specifically, we matched observed masses and retention times of
unfragmented ions to predicted masses and retention times for the
theoretical peptides from an {\it in silico} digest. We built a
statistical framework to assign protein probabilities based on these
matches, and combined the results with the ones obtained by the
typical workflow based on fragmented ions. Following this approach, we
obtained up to 7\% more protein identifications at a fixed false
discovery rate threshold of 1\% \cite{mf}.

The idea of identifying proteins based on entire peptides, rather than
peptide fragments, is not new. In fact, this was the principle
underlying peptide mass fingerprinting, one of the first methods used
to identify proteins via mass spectrometry. However, when introduced
in the 90's, mass fingerprinting could not be applied to unequivocally
identify proteins in complex biological samples. Consequently, shotgun
proteomics has gained popularity instead. Nevertheless, with great
developments in instrumentation leading to sub-ppm mass accuracies,
the possibility of using mass fingerprinting-like methods came back to
attention~\cite{iamt1}. Our study showed how improved mass accuracies
and retention time predictions enable the use of unfragmented ions
even in the context of complex peptide mixtures.

@@
